Fast Sorting on Modern Computers

### 2005-08-01 docs/tech_report/heap.tex
minor edits to fix latex errors

The most interesting thing is the relation between the predictability of the
comparisons during the destruction of the 2-, 4- and 8-heaps. Compare the
columns marked `\cc{Destruction, child comparison}' and `\cc{Destruction, child
promotion}' in Figure \ref{base and memory-tuned heapsort-1} (the 2-heap), with
the columns marked `\cc{comparison 1}' to `\cc{child promotion}' in Figure
\ref{base and memory-tuned heapsorts-2} (the 4-heap) and Figures \ref{base and
memory-tuned heapsort-3} and \ref{base and memory-tuned heapsort-4} (the
8-heap).

@@

More importantly, the accuracy of the each successive comparison increases in
the same way as selection sort (see Section \ref{selection branches logn}),
becoming more predictable as the fanout increases. For example, in the 8-heap,
shown in Figures \ref{base and memory-tuned heapsort-3} and \ref{base and
memory-tuned heapsort-4}, the first comparison is predicted correctly 50\% of
the time, while the next is predicted correctly 61\% of the time, then 71\%,
77\%, and so on in a curve, which is shown in Figure \ref{memory-tuned heapsort
curve}.

>>>>

The most interesting thing is the relation between the predictability of the
comparisons during the destruction of the 2-, 4- and 8-heaps. Compare the
columns marked `\cc{Destruction, child comparison}' and `\cc{Destruction, child
promotion}' in Figure \ref{base and memory-tuned heapsorts-1} (the 2-heap), with
the columns marked `\cc{comparison 1}' to `\cc{child promotion}' in Figure
\ref{base and memory-tuned heapsorts-2} (the 4-heap) and Figures \ref{base and
memory-tuned heapsorts-3} and \ref{base and memory-tuned heapsorts-4} (the
8-heap).

@@

More importantly, the accuracy of the each successive comparison increases in
the same way as selection sort (see Section \ref{selection branches logn}),
becoming more predictable as the fanout increases. For example, in the 8-heap,
shown in Figures \ref{base and memory-tuned heapsorts-3} and \ref{base and
memory-tuned heapsorts-4}, the first comparison is predicted correctly 50\% of
the time, while the next is predicted correctly 61\% of the time, then 71\%,
77\%, and so on in a curve, which is shown in Figure \ref{memory-tuned heapsort
curve}.

### 2005-08-01 docs/tech_report/quick.tex
minor edits to fix latex errors

The results of these tests are shown in Figure \ref{Branch prediction
performance for base quicksort without a median, with median of 3, and with
pseudo medians of 5, 7 and 9}. There is little variance in insertion sort misses
between the sorts using pseudo-medians, but the median-of-3 quicksort has
6\% fewer branches, and a corresponsing decrease in mispredictions.
%TODO to whom do we owe the idea of a pseudo median? Bently/McIlroy?
%TODO when does it cut out, and why did we choose tere (benlty?)

>>>>

The results of these tests are shown in Figure \ref{Branch prediction
performance for base quicksort without a median, with median of 3, and with
pseudo-medians of 5, 7 and 9}. There is little variance in insertion sort misses
between the sorts using pseudo-medians, but the median-of-3 quicksort has
6\% fewer branches, and a corresponsing decrease in mispredictions.
%TODO to whom do we owe the idea of a pseudo median? Bently/McIlroy?
%TODO when does it cut out, and why did we choose tere (benlty?)
